---
title: "HW3"
author: "Diwei Zhu"
date: "5/18/2022"
output: pdf_document
---


```{r}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(arrow)
library(readr)
library(gender)
library(wru)
library(lubridate)
library(ggplot2)
library(igraph)
library(ggraph)
library(tidygraph)
```

### Load data

Load the following data:
  + applications from `app_data_sample.parquet`
  + edges from `edges_sample.csv`

```{r}
data_path <- "C:/Users/admin/Documents/R projects/2022-ona-assignments/"
applications <- read_parquet(paste0(data_path,"app_data_sample.parquet"))
edges <- read_csv(paste0(data_path,"edges_sample.csv"))

applications
edges
```
## Question 1

### Get gender for examiners

We'll get gender based on the first name of the examiner, which is recorded in the field `examiner_name_first`. We'll use library `gender` for that, relying on a modified version of their own [example](https://cran.r-project.org/web/packages/gender/vignettes/predicting-gender.html).

Note that there are over 2 million records in the applications table -- that's because there are many records for each examiner, as many as the number of applications that examiner worked on during this time frame. Our first step therefore is to get all *unique* names in a separate list `examiner_names`. We will then guess gender for each one and will join this table back to the original dataset. So, let's get names without repetition:


```{r}
# install_genderdata_package() # only run this line the first time you use the package, to get data for it

# get a list of first names without repetitions
examiner_names <- applications %>% 
  distinct(examiner_name_first)

examiner_names
```

Now let's use function `gender()` as shown in the example for the package to attach a gender and probability to each name and put the results into the table `examiner_names_gender`

```{r}
# get a table of names and gender
examiner_names_gender <- examiner_names %>% 
  do(results = gender(.$examiner_name_first, method = "ssa")) %>% 
  unnest(cols = c(results), keep_empty = TRUE) %>% 
  select(
    examiner_name_first = name,
    gender,
    proportion_female
  )

examiner_names_gender
```

Finally, let's join that table back to our original applications data and discard the temporary tables we have just created to reduce clutter in our environment.

```{r}
# remove extra colums from the gender table
examiner_names_gender <- examiner_names_gender %>% 
  select(examiner_name_first, gender)

# joining gender back to the dataset
applications <- applications %>% 
  left_join(examiner_names_gender, by = "examiner_name_first")

# cleaning up
rm(examiner_names)
rm(examiner_names_gender)
gc()
```


### Guess the examiner's race

We'll now use package `wru` to estimate likely race of an examiner. Just like with gender, we'll get a list of unique names first, only now we are using surnames.


```{r}
examiner_surnames <- applications %>% 
  select(surname = examiner_name_last) %>% 
  distinct()

examiner_surnames
```
We'll follow the instructions for the package outlined here [https://github.com/kosukeimai/wru](https://github.com/kosukeimai/wru).

```{r}
examiner_race <- predict_race(voter.file = examiner_surnames, surname.only = T) %>% 
  as_tibble()

examiner_race
```

As you can see, we get probabilities across five broad US Census categories: white, black, Hispanic, Asian and other. (Some of you may correctly point out that Hispanic is not a race category in the US Census, but these are the limitations of this package.)

Our final step here is to pick the race category that has the highest probability for each last name and then join the table back to the main applications table. See this example for comparing values across columns: [https://www.tidyverse.org/blog/2020/04/dplyr-1-0-0-rowwise/](https://www.tidyverse.org/blog/2020/04/dplyr-1-0-0-rowwise/). And this one for `case_when()` function: [https://dplyr.tidyverse.org/reference/case_when.html](https://dplyr.tidyverse.org/reference/case_when.html).

```{r}
examiner_race <- examiner_race %>% 
  mutate(max_race_p = pmax(pred.asi, pred.bla, pred.his, pred.oth, pred.whi)) %>% 
  mutate(race = case_when(
    max_race_p == pred.asi ~ "Asian",
    max_race_p == pred.bla ~ "black",
    max_race_p == pred.his ~ "Hispanic",
    max_race_p == pred.oth ~ "other",
    max_race_p == pred.whi ~ "white",
    TRUE ~ NA_character_
  ))

examiner_race
```

Let's join the data back to the applications table.

```{r}
# removing extra columns
examiner_race <- examiner_race %>% 
  select(surname,race)

applications <- applications %>% 
  left_join(examiner_race, by = c("examiner_name_last" = "surname"))

rm(examiner_race)
rm(examiner_surnames)
gc()
```


## Examiner's tenure 

To figure out the timespan for which we observe each examiner in the applications data, let's find the first and the last observed date for each examiner. We'll first get examiner IDs and application dates in a separate table, for ease of manipulation. We'll keep examiner ID (the field `examiner_id`), and earliest and latest dates for each application (`filing_date` and `appl_status_date` respectively). We'll use functions in package `lubridate` to work with date and time values.

```{r}
examiner_dates <- applications %>% 
  select(examiner_id, filing_date, appl_status_date) 

examiner_dates
```

The dates look inconsistent in terms of formatting. Let's make them consistent. We'll create new variables `start_date` and `end_date`.

```{r}
examiner_dates <- examiner_dates %>% 
  mutate(start_date = ymd(filing_date), end_date = as_date(dmy_hms(appl_status_date)))
```

Let's now identify the earliest and the latest date for each examiner and calculate the difference in days, which is their tenure in the organization.

```{r}
examiner_dates <- examiner_dates %>% 
  group_by(examiner_id) %>% 
  summarise(
    earliest_date = min(start_date, na.rm = TRUE), 
    latest_date = max(end_date, na.rm = TRUE),
    tenure_days = interval(earliest_date, latest_date) %/% days(1)
    ) %>% 
  filter(year(latest_date)<2018)

examiner_dates
```

Joining back to the applications data.

```{r}
applications <- applications %>% 
  left_join(examiner_dates, by = "examiner_id")

rm(examiner_dates)
gc()
```



## Question2


```{r}
# we pick work group 163 and 172

w164 <- subset(applications, grepl("^164", applications$examiner_art_unit))
w164$gender <- factor(w164$gender)
w164$race <- factor(w164$race)
w241 <- subset(applications, grepl("^241", applications$examiner_art_unit))
w241$gender <- factor(w241$gender)
w241$race <- factor(w241$race)

# summary for group 164
summary(w164$gender)
summary(w164$race)
summary(w164$tenure_days)

# summry for group 241
summary(w241$gender)
summary(w241$race)
summary(w241$tenure_days)
```

Race and gender distribution for work group 164 and 241 respectively
```{r}
# merge 
w164$workgroup <- c('164')
w241$workgroup <- c('241')

merged = union(x = w164,y = w241)
```

Gender distribution in races for work group 164
```{r}
toPlot<-w164%>%
  group_by(gender, race)%>%
  summarise(n = n())%>%
  group_by(race)%>%
  mutate(prop = n/sum(n))

ggplot(data = toPlot, aes(gender, prop, fill = race)) + 
  geom_col() + 
  facet_grid(~race)+
  scale_fill_manual(values = c("lightyellow3","lightsalmon3", "wheat3","white"))
```

Gender distribution in races for work group 241
```{r}
toPlot<-w241 %>%
  group_by(gender, race)%>%
  summarise(n = n())%>%
  group_by(race)%>%
  mutate(prop = n/sum(n))

ggplot(data = toPlot, aes(gender, prop, fill = race)) + 
  geom_col() + 
  facet_grid(~race)+
  scale_fill_manual(values = c("lightyellow3","lightsalmon3", "wheat3","white"))
```


## Question 3
Create node lists for eacch work group
```{r}
# join selected work groups with edges list
edges <- drop_na(edges, ego_examiner_id)
edges <-drop_na(edges, alter_examiner_id)
w164_2 <- inner_join(w164, edges, by = "application_number", copy = FALSE) 
w241_2 <- inner_join(w241, edges, by = "application_number", copy = FALSE) 

# nodes dataframe of work groups and merge them
w164_nodes1 <- w164_2 %>% 
  distinct(ego_examiner_id) %>%
  rename(ID = ego_examiner_id)

w164_nodes2 <- w164_2 %>% 
  distinct(alter_examiner_id) %>%
  rename(ID = alter_examiner_id)

w241_nodes1 <- w241_2 %>% 
  distinct(ego_examiner_id) %>%
  rename(ID = ego_examiner_id)

w241_nodes2 <- w241_2 %>% 
  distinct(alter_examiner_id) %>%
  rename(ID = alter_examiner_id)

# merge the two dataframes for each work goup
w164_nodes <- union_all(w164_nodes1, w164_nodes2)
w241_nodes <- union_all(w241_nodes1, w241_nodes2)

w164_nodes <- unique(w164_nodes)
w241_nodes <- unique(w241_nodes)

head(w164_nodes, 5)
```

Create final edge list
```{r}
w164_edges <- w164_2 %>% 
  select(ego_examiner_id, alter_examiner_id)

w241_edges <- w241_2 %>% 
  select(ego_examiner_id, alter_examiner_id)

head(w164_edges, 5)
```


```{r}
g_w164 <- graph_from_data_frame(w164_edges, directed=FALSE)
g_w241 <- graph_from_data_frame(w241_edges, directed=FALSE)
```

Plot vertex graph for work group 164
```{r}
plot(g_w164, layout=layout.fruchterman.reingold,
    vertex.size = 5,
    vertex.label = NA,
    vertex.color = "palevioletred2")
```

Plot vertex graph for work group 241
```{r}
plot(g_w241, layout=layout.fruchterman.reingold,
    vertex.size = 5,
    vertex.label = NA,
    vertex.color = "darkslategray3")
```
Calculate centralities
```{r}
# betweenness
bc_w164 <- sort(betweenness(g_w164), decreasing = TRUE)
bc_w241 <- sort(betweenness(g_w241), decreasing = TRUE)

# degree
dg_w164 <- sort(degree(g_w164), decreasing = TRUE)
dg_w241 <- sort(degree(g_w241), decreasing = TRUE)

# closeness
cc_w164 <- sort(closeness(g_w164), decreasing = TRUE)
cc_w241 <- sort(closeness(g_w241), decreasing = TRUE)

print("top 5 of betwenness centrality for work group 164")
print(head(bc_w164,5))
print("top 5 of betwenness centrality for work group 241")
print(head(bc_w241,5))
print("top 5 of degree centrality for work group 164")
print(head(dg_w164,5))
print("top 5 of degree centrality for work group 241")
print(head(dg_w241,5))
print("top 5 of closeness centrality for work group 164")
print(head(cc_w164,15))
print("top 5 of closeness centrality for work group 241")
print(head(cc_w241,15))
```

### Pick measures of centrality
I would pick betweenness centrality and degree centrality from the three centralities I calculated above, because in the USPTO case: (1) how many contacts does one examiner have (degree centrality) is a useful information, and (2) the transfer of information among all examiners are important, making the examiner who on the shortest paths between other examiners more important. Additionally, because for work group 241, there are many examiners who are of the highest closeness centrality, the advantage of having the highest closeness centrality is not very strong in my specific case. 




